---
title: "Does ARK (still) Generate Alpha?"
author: "Martin Geissmann"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: pdf_document
abstract: "Ark Investment has become the largest actively managed ETF provider with innovation/technology-themed funds that delivered great performance. The question one should always ask is if putting your money in any fund will deliver exceptional returns (alpha). We find that all ARK funds delivered a mix of alpha and beta. Above average beta was responsible for the exceptional returns over the past years."
---

```{r init, include=FALSE, echo = F, message=F, warning=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
options(knitr.kable.NA = "")
```

```{r}
library(tidyverse)
library(tiqyquant)
library(lubridate)
```

```{r}
yahoo_monthly_rets <- function(symbol) {
  tq_get(symbol) %>%
    group_by(m = floor_date(date, "months")) %>%
    filter(date == max(date)) %>% 
    ungroup() %>% 
    select(-m) %>% 
    transmute(symbol, date, r = adjusted/lag(adjusted)-1) %>% 
    filter(!is.na(r))
}

lookup <- read_lines("SNPE   Xtrackers S&P 500 ESG ETF
                      ESGU   iShares ESG Aware MSCI USA ETF
                      SUSA   iShares MSCI USA ESG Select ETF 
                      SUSL   iShares ESG MSCI USA Leaders ETF
                      USXF   iShares ESG Advanced MSCI USA ETF 
                      IQSU   IQ Candriam ESG US Equity ETF
                      NUEM   Nuveen ESG Emerging Markets Equity ETF
                      EMSG   Xtrackers MSCI Emerging Markets ESG Leaders Equity ETF
                      RESE   WisdomTree Emerging Markets ESG Fund
                      ESGD   iShares ESG Aware MSCI EAFE ETF
                      EASG   Xtrackers MSCI EAFE ESG Leaders Equity ETF
                      DMXF   iShares ESG Advanced MSCI EAFE Index ETF")

lookup <- tibble(raw = lookup %>% 
         str_trim()) %>% 
  separate(col = raw, into = c("symbol", "name"), sep = "  ") %>% 
  mutate(name = str_trim(name))
```

```{r}
lookup %>% 
  kable(caption = "ETFs in focus") %>% 
  kable_styling(latex_options = "HOLD_position")
```

```{r, eval = F}
# SNPE   Xtrackers S&P 500 ESG ETF - 2019-07-31
# Benchmark: SPY
SNPE <- yahoo_monthly_rets("SNPE")
SPY  <- yahoo_monthly_rets("SPY")

SNPE <- SNPE %>% 
  inner_join(SPY %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)

# ESGU   iShares ESG Aware MSCI USA ETF - 2017-10-31
# SUSA   iShares MSCI USA ESG Select ETF - 2017-10-31
# SUSL   iShares ESG MSCI USA Leaders ETF - 2019-06-28
# USXF   iShares ESG Advanced MSCI USA ETF - 2020-07-31
# Benchmark: PBUS
ESGU <- yahoo_monthly_rets("ESGU")
SUSA <- yahoo_monthly_rets("SUSA")
SUSL <- yahoo_monthly_rets("SUSL")
USXF <- yahoo_monthly_rets("USXF")
PBUS <- yahoo_monthly_rets("PBUS")

ESGU <- ESGU %>% 
  inner_join(PBUS %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
SUSA <- SUSA %>% 
  inner_join(PBUS %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
SUSL <- SUSL %>% 
  inner_join(PBUS %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
USXF <- USXF %>% 
  inner_join(PBUS %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)


# IQSU   IQ Candriam ESG US Equity ETF - 2020-01-31
# Benchmark: VTI
IQSU <- yahoo_monthly_rets("IQSU")
VTI  <- yahoo_monthly_rets("VTI")

IQSU <- IQSU %>% 
  inner_join(VTI %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)

# NUEM   Nuveen ESG Emerging Markets Equity ETF - 2017-07-31
# EMSG   Xtrackers MSCI Emerging Markets ESG Leaders Equity ETF - 2019-01-31
# RESE   WisdomTree Emerging Markets ESG Fund - 2020-04-30
# Benchmark: EEM
NUEM <- yahoo_monthly_rets("NUEM")
EMSG <- yahoo_monthly_rets("EMSG")
RESE <- yahoo_monthly_rets("RESE")
EEM <- yahoo_monthly_rets("EEM")

NUEM <- NUEM %>% 
  inner_join(EEM %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
EMSG <- EMSG %>% 
  inner_join(EEM %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
RESE <- RESE %>% 
  inner_join(EEM %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)

# ESGD   iShares ESG Aware MSCI EAFE ETF - 2016-09-30
# EASG   Xtrackers MSCI EAFE ESG Leaders Equity ETF - 2018-12-31
# DMXF   iShares ESG Advanced MSCI EAFE Index ETF - 2020-07-31
# Benchmark: IEFA
ESGD <- yahoo_monthly_rets("ESGD")
EASG <- yahoo_monthly_rets("EASG")
DMXF <- yahoo_monthly_rets("DMXF")
IEFA <- yahoo_monthly_rets("IEFA")

ESGD <- ESGD %>% 
  inner_join(IEFA %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
EASG <- EASG %>% 
  inner_join(IEFA %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)
DMXF <- DMXF %>% 
  inner_join(IEFA %>% 
               select(date, bm = r),
             by = "date") %>% 
  mutate(r_excess = r - bm)


SNPE %>% saveRDS("data/SNPE")
ESGU %>% saveRDS("data/ESGU")
SUSA %>% saveRDS("data/SUSA")
SUSL %>% saveRDS("data/SUSL")
USXF %>% saveRDS("data/USXF")
IQSU %>% saveRDS("data/IQSU")
NUEM %>% saveRDS("data/NUEM")
EMSG %>% saveRDS("data/EMSG")
RESE %>% saveRDS("data/RESE")
ESGD %>% saveRDS("data/ESGD")
EASG %>% saveRDS("data/EASG")
DMXF %>% saveRDS("data/DMXF")
```

```{r, echo = F}
SNPE <- readRDS("data/SNPE")
ESGU <- readRDS("data/ESGU")
SUSA <- readRDS("data/SUSA")
SUSL <- readRDS("data/SUSL")
USXF <- readRDS("data/USXF")
IQSU <- readRDS("data/IQSU")
NUEM <- readRDS("data/NUEM")
EMSG <- readRDS("data/EMSG")
RESE <- readRDS("data/RESE")
ESGD <- readRDS("data/ESGD")
EASG <- readRDS("data/EASG")
DMXF <- readRDS("data/DMXF")
```

```{r}
df <- bind_rows(SNPE,
                ESGU,
                SUSA,
                SUSL,
                USXF,
                IQSU,
                NUEM,
                EMSG,
                RESE,
                ESGD,
                EASG,
                DMXF) %>% 
  left_join(lookup,
            by = "symbol")

df_r <- full_join(SNPE %>% select(date, SNPE = r),
          ESGU %>% select(date, ESGU = r),
          by = "date") %>% 
  full_join(SUSA %>% select(date, SUSA = r),
            by = "date") %>% 
  full_join(SUSL %>% select(date, SUSL = r),
            by = "date") %>% 
  full_join(USXF %>% select(date, USXF = r),
            by = "date") %>% 
  full_join(IQSU %>% select(date, IQSU = r),
            by = "date") %>% 
  full_join(NUEM %>% select(date, NUEM = r),
            by = "date") %>% 
  full_join(EMSG %>% select(date, EMSG = r),
            by = "date") %>% 
  full_join(RESE %>% select(date, RESE = r),
            by = "date") %>% 
  full_join(ESGD %>% select(date, ESGD = r),
            by = "date") %>% 
  full_join(EASG %>% select(date, EASG = r),
            by = "date") %>% 
  full_join(DMXF %>% select(date, DMXF = r),
            by = "date") %>% 
  arrange(date) %>% 
  pivot_longer(-date, values_to = "r", names_to = "symbol") %>% 
  arrange(symbol, date)

df_excess_r <- full_join(SNPE %>% select(date, SNPE = r_excess),
                         ESGU %>% select(date, ESGU = r_excess),
                         by = "date") %>% 
  full_join(SUSA %>% select(date, SUSA = r_excess),
            by = "date") %>% 
  full_join(SUSL %>% select(date, SUSL = r_excess),
            by = "date") %>% 
  full_join(USXF %>% select(date, USXF = r_excess),
            by = "date") %>% 
  full_join(IQSU %>% select(date, IQSU = r_excess),
            by = "date") %>% 
  full_join(NUEM %>% select(date, NUEM = r_excess),
            by = "date") %>% 
  full_join(EMSG %>% select(date, EMSG = r_excess),
            by = "date") %>% 
  full_join(RESE %>% select(date, RESE = r_excess),
            by = "date") %>% 
  full_join(ESGD %>% select(date, ESGD = r_excess),
            by = "date") %>% 
  full_join(EASG %>% select(date, EASG = r_excess),
            by = "date") %>% 
  full_join(DMXF %>% select(date, DMXF = r_excess),
            by = "date") %>% 
  arrange(date) %>% 
  pivot_longer(-date, values_to = "r", names_to = "symbol") %>% 
  arrange(symbol, date)
```

```{r}
df_r %>% 
  filter(date >= as.Date("2017-06-01")) %>% 
  ggplot(aes(x = date, y = r, fill = symbol)) +
  geom_col(width = 25, position = "dodge", ) +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_minor_breaks = "months") +
  labs(x = "", y = "", fill = "", title = "Absolute Monthly Return of the ESG ETFs") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
df_excess_r %>% 
  filter(date >= as.Date("2017-06-01")) %>% 
  ggplot(aes(x = date, y = r, fill = symbol)) +
  geom_col(width = 25, position = "dodge", ) +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_minor_breaks = "months") +
  labs(x = "", y = "", fill = "", title = "Simple Excess Return of the ESG ETFs over their Benchmarks") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_excess_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(!str_detect(name, "Emerging"), 
         !str_detect(name, "EAFE")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Excess Returns since ", format(d, "%m.%Y"), " - USA"),
                     mar = c(2, 0, 1, 0))
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(!str_detect(name, "Emerging"), 
         !str_detect(name, "EAFE")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Absolute Returns since ", format(d, "%m.%Y"), " - USA"),
                     mar = c(2, 0, 1, 0))
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_excess_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(str_detect(name, "EAFE")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Excess Returns since ", format(d, "%m.%Y"), " - EAFE"),
                     mar = c(2, 0, 1, 0))
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(str_detect(name, "EAFE")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Absolute Returns since ", format(d, "%m.%Y"), " - EAFE"),
                     mar = c(2, 0, 1, 0))
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_excess_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(str_detect(name, "Emerging")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Excess Returns since ", format(d, "%m.%Y"), " - Emerging Markets"),
                     mar = c(2, 0, 1, 0))
```

```{r, fig.height=7}
d <- as.Date("2020-07-01")
par(xpd=TRUE)
df_r %>% 
  filter(date >= d) %>%
  arrange(date) %>% 
  left_join(lookup, by = "symbol") %>% 
  filter(str_detect(name, "Emerging")) %>% 
  mutate(name = gsub('(.{18})\\s(.*)', '\\1\n\\2', name)) %>% 
  select(-symbol) %>% 
  group_by(name) %>% 
  filter(all(!is.na(r))) %>% 
  pivot_wider(names_from = name, values_from = r) %>% 
  select(-date) %>% 
  cor() %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = paste0("Correlations of Absolute Returns since ", format(d, "%m.%Y"), " - Emerging Markets"),
                     mar = c(2, 0, 1, 0))
```

Why are the correlations in excess returns so low or even negative? It seems that there is not one *ESG factor* driving performance. The ESG selection of stocks can randomly perform better or worse.

Going back to the chart over time, we saw that the excess returns seem to depend on the underlying abolute performance. This might be an indication that the ESG products have an above average market beta.

# Beta Adjustment

```{r}
# Rolling Alpha ----
regr_fun_beta <- function(d, col = "r", bm_col = "bm") {
  # https://www.r-bloggers.com/2017/04/tidyquant-0-5-0-select-rollapply-and-quandl/
  # https://stackoverflow.com/questions/55877110/pass-dynamically-variable-names-in-lm-formula-inside-a-function
  coefs <- lm(d[,col] ~ d[,bm_col]) %>%
    coef()
  beta = coefs[2]
  return(beta)
}

beta <- df %>% 
  group_by(name) %>% 
  tq_mutate(mutate_fun = rollapply,
            width      = 12,
            FUN        = regr_fun_beta,
            by.column  = FALSE,
            col = "r",
            col_rename = "beta") %>% 
  ungroup()
```

```{r}
longest_etfs <- df %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  tail(6)

beta %>% 
  filter(name %in% longest_etfs$name) %>% 
  ggplot(aes(x = date, y = beta, color = name)) +
  geom_line() +
  geom_hline(yintercept = 1, color = "red", linetype = 2) +
  labs(x = "", y = "beta on benchmark", color = "") +
  guides(color=guide_legend(ncol=2,byrow=TRUE)) +
  theme_minimal() +
  theme(legend.position = "top") 
```

```{r}
df_excess_r_adjusted <- beta %>% 
  mutate(r_excess_adjusted = r - beta*bm) %>% 
  select(date, symbol, name, r = r_excess_adjusted) %>% 
  arrange(symbol, date)

df_excess_r_adjusted %>% 
  filter(date >= as.Date("2017-06-01")) %>% 
  ggplot(aes(x = date, y = r, fill = symbol)) +
  geom_col(width = 25, position = "dodge", ) +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_minor_breaks = "months") +
  labs(x = "", y = "", fill = "", title = "Absolute Monthly Return of the ESG ETFs") +
  theme_minimal() +
  theme(legend.position = "none")
```

We can conclude that market beta is not a main driver.