---
title: "Backtesting Trading Strategies in R"
author: "Martin Geissmann"
date: "`r Sys.setlocale('LC_TIME', 'en_US.UTF-8'); format(Sys.Date(), '%d %b %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: true
# abstract: ""
---

```{r, include=F}
library(knitr)
library(kableExtra)

library(tidyverse)
```

# Introduction

**Backtesting is simulating a systematic trading strategy on past data.**

For example:

* Always invest in equal amounts in SPY, QQQ, and TLT. Rebalance annually.
* Invest in IWM if above 200-day moving average, otherwise AGG.
* Invest in the tangency portfolio (Markowitz' model) using all S&P 500 stocks.
* Invest in the 10 stocks with the highest retail trading volume (alternative data)

In general, a backtest is the result of two steps:

1. Signal \rightarrow Theoretical portfolio weights
2. Theoretical portfolio weights \rightarrow Actual weights and portfolio return 

The second step is subject to assumptions, e.g. the rebalance frequency, fees, slippage etc.

What are our options for backtesting trading strategies in R?

* "Basic" R (xts or tibble/dplyr)
* Easy backtesting functions/packages
* Accounting backtesting packages

## How to find R packages

Packages by name: https://cran.r-project.org/web/packages/available_packages_by_name.html

Search for packages using `packagefinder::findPackage`

```{r, eval = F}
library(packagefinder)

findPackage("backtest")
```

## Packages

```{r, echo=F}
comp <- readxl::read_excel("package-comparison.xlsx")

comp %>% 
  select(-link) %>% 
  mutate(url = "link") %>% 
  kable(caption = "Packages",
        booktabs = T, 
        linesep = "") %>% 
  kable_styling(latex_options = c("hold_position", "striped", "scale_down")) %>% 
  column_spec(1, width = "7em") %>% 
  column_spec(2, width = "40em") %>% 
  column_spec(3, width = "3em", link = comp$link)
```

# dplyr

## SMA Rotation Strategy (Risk On/Off)

*Rule*: If the S&P 500 is trading above its 200-days SMA, then invest in stocks. If it's below, then rotate into gold.

```{r, message=F, eval = F}
library(tidyverse)
library(lubridate)
library(tidyquant)
library(RcppRoll)
library(scales)
```

Get data.

```{r, eval = F}
spy <- tq_get("SPY", from = as.Date("2005-01-01"))
gld <- tq_get("GLD", from = as.Date("2005-01-01"))
```

```{r, eval = F}
spy_gld <- full_join(spy %>% 
                       select(date, spy.cl = close, spy.adj = adjusted),
                     gld %>% 
                       select(date, gld.adj = adjusted),
                     by = "date") %>% 
  arrange(date) %>% 
  mutate(spy.r = spy.adj/lag(spy.adj)-1,
         gld.r = gld.adj/lag(gld.adj)-1)
```

```{r, eval = F}
strategy <- spy_gld %>% 
  mutate(SMA = roll_mean(spy.cl, n = 200, align = "right", fill = NA_real_),
         SMA_signal = lag(spy.cl >= SMA)) %>% 
  filter(!is.na(SMA_signal)) %>% 
  mutate(strategy.r = ifelse(SMA_signal, spy.r, gld.r))
```

```{r, eval = F}
strategy_long <- strategy %>% 
  select(date, spy.r, gld.r, strategy.r) %>% 
  pivot_longer(-date, values_to = "r") %>% 
  arrange(name, date) %>% 
  group_by(name) %>% 
  # cumulative performance
  mutate(p = cumprod(1+r)-1) %>% 
  group_by(series = str_to_upper(str_remove(name, ".r$")))
```

```{r, eval = F}
strategy_long %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_color_brewer(type = "qual", palette = 3) +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b %y") +
  labs(x = "", y = "", color = "") +
  theme_minimal() +
  theme(legend.position = "top")
```

```{r, eval = F}
strategy_long %>% 
  # annualized performance (PerformanceAnalytics, tidyquant)
  tq_performance(Ra = r, performance_fun = table.AnnualizedReturns, scale = 252)
```

```{r, eval = F}
strategy_long %>% 
  # calculate drawdowns
  mutate(dd = (1+p)/cummax(1+p)-1) %>% 
  ggplot(aes(x = date, y = dd, color = series)) +
  geom_line() +
  scale_color_brewer(type = "qual", palette = 3) +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_labels = "%b %y") +
  labs(x = "", y = "", color = "") +
  theme_minimal() +
  theme(legend.position = "top")
```

# portfolioBacktest

- requires xts data
- ideal of testing on several subsamples of data at once

```{r, eval = F}
library(portfolioBacktest)

data("dataset10")

dataset10$`dataset 1`$adjusted

# defining the portfolio (equal weighting)
my_portfolio <- function(dataset, ...) {
  prices <- dataset$adjusted
  print(nrow(prices))
  N <- ncol(prices)
  return(rep(1/N, N))
}

bt <- portfolioBacktest(my_portfolio, list(dataset10$`dataset 1`))

backtestSummary(bt)$performance

my_portfolio_sma <- function(dataset, ...) {
  prices <- dataset$adjusted
  print(prices)
  
    N <- ncol(prices)
  return(rep(1/N, N))
}

bt <- portfolioBacktest(portfolio_funs = my_portfolio_sma(), dataset_list = dataset10)

```

# AssetAllocation

```{r, eval = F}
library(AssetAllocation)
library(PerformanceAnalytics)
names(basic_asset_alloc)
```

```{r, eval = F}
basic_asset_alloc$all_weather

ivy_port_weights()
```

```{r, eval = F}
## Example 1: backtesting one of the asset allocations in the package
strat_permanent <- basic_asset_alloc$permanent
strat_fancy <- list(name = "Fancy",
                    tickers = c("SPY", "TLT"),
                    default_weights = c(1, 0),
                    rebalance_frequency = "month",
                    portfolio_rule_fn)

# test using the data set provided in the package
bt_strat_permanent <- backtest_allocation(strat_permanent, ETFs_daily)

# plot cumulative returns
chart.CumReturns(bt_strat_permanent$returns,
                 main = paste0("Cumulative returns of the ",
                               bt_strat_permanent$strat$name,
                               " portfolio"),
                 ylab = "Cumulative returns")
```

Only supports constant weights for now.
