---
title: "How bad did the Fama French Factors perform in the past decade?"
author: "Martin Geissmann"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  pdf_document:
    number_sections: true
abstract: "Value was beaten by Growth. The Small Cap premium vanished in a decade where big tech won. But how bad has it actually been? And what's the picture outside of the US?"
---

# Data source

https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html

The factors are abbreviated as follows: *Mkt-RF*: Market excess return, *SMB*: Small-Minus-Big (Size), *HML*: High-Minus-Low (Book-to-Market Value), *RMW*: Robust-Minus-Weak (Profitability), *CMA*: Conservative-Minus-Aggressive (Investment), *WML*: Winners-Minus-Losers (Momentum)

```{r init, include=FALSE, echo = F, message=F, warning=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
options(knitr.kable.NA = "")
```

```{r}
library(tidyverse)
library(lubridate)
library(scales)
```

```{r}
ff_5 <- read_csv("data/F-F_Research_Data_5_Factors_2x3.csv", skip = 3,
                 na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(ff_5$date)))

ff_5 <- ff_5 %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date)


ff_mom <- read_csv("data/F-F_Momentum_Factor.CSV", skip = 13,
                   na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(ff_mom$date)))

ff_mom <- ff_mom %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date) %>% 
  filter(date >= min(ff_5$date))

ff_5_mom <- bind_rows(ff_5,
                      ff_mom) %>% 
  mutate(series = if_else(series == "Mom", "WML", series))
```

```{r}
europe_5 <- read_csv("data/Europe_5_Factors.csv", skip = 6,
                     na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(europe_5$date)))

europe_5 <- europe_5 %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date)


europe_mom <- read_csv("data/Europe_MOM_Factor.csv", skip = 6,
                   na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(europe_mom$date)))

europe_mom <- europe_mom %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date) %>% 
  filter(date >= min(ff_5$date))

europe_5_mom <- bind_rows(europe_5,
                      europe_mom)
```

```{r}
japan_5 <- read_csv("data/Japan_5_Factors.csv", skip = 6,
                    na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(japan_5$date)))

japan_5 <- japan_5 %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date)


japan_mom <- read_csv("data/Japan_MOM_Factor.csv", skip = 6,
                   na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(japan_mom$date)))

japan_mom <- japan_mom %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date) %>% 
  filter(date >= min(ff_5$date))

japan_5_mom <- bind_rows(japan_5,
                          japan_mom)
```

```{r}
asiapacific_5 <- read_csv("data/Asia_Pacific_ex_Japan_5_Factors.csv", skip = 6,
                    na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(asiapacific_5$date)))

asiapacific_5 <- asiapacific_5 %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date)


asiapacific_mom <- read_csv("data/Asia_Pacific_ex_Japan_MOM_Factor.csv", skip = 6,
                   na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(asiapacific_mom$date)))

asiapacific_mom <- asiapacific_mom %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date) %>% 
  filter(date >= min(ff_5$date))

asiapacific_5_mom <- bind_rows(asiapacific_5,
                          asiapacific_mom)
```

```{r}
emerging_5 <- read_csv("data/Emerging_5_Factors.csv", skip = 6,
                    na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(emerging_5$date)))

emerging_5 <- emerging_5 %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date)


emerging_mom <- read_csv("data/Emerging_MOM_Factor.csv", skip = 6,
                   na = c("", "NA", "-99.99", "-999")) %>% 
  mutate(date = ymd(paste0(X1, "01")))

last_row <- min(which(is.na(emerging_mom$date)))

emerging_mom <- emerging_mom %>% 
  filter(row_number() < last_row) %>% 
  select(-X1) %>% 
  pivot_longer(-date, names_to = "series", values_to = "r") %>% 
  mutate(r = as.double(r)/100) %>% 
  arrange(series, date) %>% 
  filter(date >= min(ff_5$date))

emerging_5_mom <- bind_rows(emerging_5,
                          emerging_mom)
```

# Performance over time

The performance of the FF factors was poor over the past decade 2010 - 2019. We observed similarly poor performances in the 1990s. In both of these decades we had strong stock market returns.

The pattern is especially observable in the Standard FF factors which include US stock (NYSE, AMEX, and NASDAQ).

The factors are abbreviated as follows:

* *Mkt-RF*: Market excess return
* *SMB*: Small-Minus-Big (Size)
* *HML*: High-Minus-Low (Book-to-Market Value)
* *RMW*: Robust-Minus-Weak (Profitability) 
* *CMA*: Conservative-Minus-Aggressive (Investment)
* *WML*: Winners-Minus-Losers (Momentum)

## Standard Fama French Factors

```{r, fig.height=9}
df <- ff_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12 | dec == "2020s") %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_breaks = "2 years", date_labels = "%y") +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(name = "Set3", n = 10))) +
  labs(x = "", y = "", color = "",
       title = "Fama French Factors (Standard)") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~dec, scales = "free_x")
```

## Europe Fama French Factors

```{r}
df <- europe_5_mom %>% 
  filter(!is.na(r)) %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(floor_date(from, "year"), " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_breaks = "2 years", date_labels = "%y") +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(name = "Set3", n = 10))) +
  labs(x = "", y = "", color = "",
       title = "Fama French Factors (Europe)") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~dec, scales = "free_x")
```

## Japan Fama French Factors

```{r}
df <- japan_5_mom %>% 
  filter(!is.na(r)) %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(floor_date(from, "year"), " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_breaks = "2 years", date_labels = "%y") +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(name = "Set3", n = 10))) +
  labs(x = "", y = "", color = "",
       title = "Fama French Factors (Japan)") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~dec, scales = "free_x")
```

## Asia Pacific ex Japan Fama French Factors

```{r}
df <- asiapacific_5_mom %>% 
  filter(!is.na(r)) %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(floor_date(from, "year"), " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_breaks = "2 years", date_labels = "%y") +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(name = "Set3", n = 10))) +
  labs(x = "", y = "", color = "",
       title = "Fama French Factors (Asia Pacific ex Japan)") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~dec, scales = "free_x")
```

## Emerging Markets Fama French Factors

```{r, fig.height=9}
df <- emerging_5_mom %>% 
  filter(!is.na(r)) %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(floor_date(from, "year"), " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df %>% 
  ggplot(aes(x = date, y = p, color = series)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  scale_x_date(date_breaks = "2 years", date_labels = "%y") +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(name = "Set3", n = 10))) +
  labs(x = "", y = "", color = "",
       title = "Fama French Factors (Emerging Markets)") +
  theme_minimal() +
  theme(legend.position = "top") +
  facet_wrap(~dec, scales = "free_x")
```

# Correlations

## Standard Fama French Factors

```{r}
df <- ff_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df_cor <- df %>% 
  ungroup() %>% 
  select(date, series, r) %>% 
  pivot_wider(names_from = series, values_from = r) %>% 
  select(-date) %>%
  select(`_Mkt-RF`, everything()) %>% 
  cor()

df_cor %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = "Correlations of Standard Monthly FF-Factors 1970 - 2021",
                     mar = c(2, 0, 1, 0))
```

## Europe Fama French Factors

```{r}
df <- europe_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df_cor <- df %>% 
  ungroup() %>% 
  select(date, series, r) %>% 
  pivot_wider(names_from = series, values_from = r) %>% 
  select(-date) %>%
  select(`_Mkt-RF`, everything()) %>% 
  filter(!is.na(WML)) %>% 
  cor()

df_cor %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = "Correlations of European Monthly FF-Factors 1990 - 2021",
                     mar = c(2, 0, 1, 0))
```

## Japan Fama French Factors

```{r}
df <- japan_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df_cor <- df %>% 
  ungroup() %>% 
  select(date, series, r) %>% 
  pivot_wider(names_from = series, values_from = r) %>% 
  select(-date) %>%
  select(`_Mkt-RF`, everything()) %>% 
  filter(!is.na(WML)) %>% 
  cor()

df_cor %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = "Correlations of Japanese Monthly FF-Factors 1990 - 2021",
                     mar = c(2, 0, 1, 0))
```

## Asia Pacific ex Japan Fama French Factors

```{r}
df <- asiapacific_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df_cor <- df %>% 
  ungroup() %>% 
  select(date, series, r) %>% 
  pivot_wider(names_from = series, values_from = r) %>% 
  select(-date) %>%
  select(`_Mkt-RF`, everything()) %>% 
  filter(!is.na(WML)) %>% 
  cor()

df_cor %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = "Correlations of Asia Pacific exJap Monthly FF-Factors 1990 - 2021",
                     mar = c(2, 0, 1, 0))
```

## Emerging Markets Fama French Factors

```{r}
df <- emerging_5_mom %>% 
  group_by(decade = floor_date(date, "10 years"), series) %>% 
  mutate(from = min(date), to = max(date),
         p = cumprod(1+r)-1,
         from_to = paste0(from, " - ", to),
         dec = paste0(format(from, "%Y"), "s")) %>% 
  filter(n() > 8*12) %>% 
  mutate(series = ifelse(series == "Mkt-RF", "_Mkt-RF", series)) %>% 
  filter(series != "RF")

df_cor <- df %>% 
  ungroup() %>% 
  select(date, series, r) %>% 
  pivot_wider(names_from = series, values_from = r) %>% 
  select(-date) %>%
  select(`_Mkt-RF`, everything()) %>% 
  filter(!is.na(WML), !is.na(CMA)) %>% 
  cor()

df_cor %>% 
  corrplot::corrplot(diag = T, type = "lower", method = "color", addCoef.col = "black",
                     tl.cex=0.9, 
                     title = "Correlations of Asia Pacific exJap Monthly FF-Factors 1990 - 2021",
                     mar = c(2, 0, 1, 0))
```

